---
title: "DATA 621 Homework #1"
author: "Calvin, Juanelle, Kevin, Ravi, Sudhan"
date: "9/25/2019"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
        theme: lumen
        number_sections: TRUE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Exploration

Load and do initial review of the data.  Need to address missing data.

```{r message = FALSE}
library(tidyverse)
library(funModeling)

gh <- "https://raw.githubusercontent.com/kecbenson/DATA621/master/HW1/"
file_train <- paste0(gh, "moneyball-training-data.csv")
file_test <- paste0(gh, "moneyball-evaluation-data.csv")

df_train <- read_csv(file_train)
df_test <- read_csv(file_test)

head(df_train)
str(df_train)
summary(df_train)
attach(df_train)
```

Check distributions of the data; skewness may need to be dealt with transformations. 

```{r cache = TRUE}
hist(df_train$TARGET_WINS)
#pairs(df_train)

for (j in 2:ncol(df_train)) {
    hist(df_train[[j]], main = paste0("Histogram of ", colnames(df_train)[j]),
         xlab = colnames(df_train)[j], freq = FALSE)
    minval <- min(df_train[[j]], na.rm = TRUE)
    maxval <- max(df_train[[j]], na.rm = TRUE)
    meanval <- mean(df_train[[j]], na.rm = TRUE)
    sdval <- sd(df_train[[j]], na.rm = TRUE)
    grid <- minval:maxval
    lines(grid, dnorm(grid, mean = meanval, sd = sdval), lty = 3)
    
}
```

Review pair-wise relationships between predictor variables and target variable.

```{r cache = TRUE}
# batting variables (hits through home runs)
pairs(df_train[2:6])
# batting variables (walks, strikeouts, hit by pitch)
pairs(df_train[c(2, 7:8, 11)])
# baserun and fielding variables
pairs(df_train[c(2, 9:10, 16:17)])
# pitching variables
pairs(df_train[c(2, 12:15)])
```


# Data Preparation

```{r}
# trigger a dummy variable if NA is present
df_train$HBP_missing <- ifelse(is.na(df_train$TEAM_BATTING_HBP), 1, 0)
# imputing mean value as a replacement to NA
df_train$TEAM_BATTING_HBP[is.na(df_train$TEAM_BATTING_HBP)] <- mean(df_train$TEAM_BATTING_HBP, na.rm=TRUE)
```

```{r}
# trigger a dummy variable if NA is present
df_train$CS_missing <- ifelse(is.na(df_train$TEAM_BASERUN_CS), 1, 0)
# imputing mean value as a replacement to NA
df_train$TEAM_BASERUN_CS[is.na(df_train$TEAM_BASERUN_CS)] <- mean(df_train$TEAM_BASERUN_CS, na.rm=TRUE)
```

```{r}
# trigger a dummy variable if NA is present
df_train$DP_missing <- ifelse(is.na(df_train$TEAM_FIELDING_DP), 1, 0)
# imputing mean value as a replacement to NA
df_train$TEAM_FIELDING_DP[is.na(df_train$TEAM_FIELDING_DP)] <- mean(df_train$TEAM_FIELDING_DP, na.rm=TRUE)
```

```{r}
# imputing mean value as a replacement to NA
df_train$TEAM_BASERUN_SB[is.na(df_train$TEAM_BASERUN_SB)] <- mean(df_train$TEAM_BASERUN_SB, na.rm=TRUE)
# imputing mean value as a replacement to NA
df_train$TEAM_BATTING_SO[is.na(df_train$TEAM_BATTING_SO)] <- mean(df_train$TEAM_BATTING_SO, na.rm=TRUE)
# imputing mean value as a replacement to NA
df_train$TEAM_PITCHING_SO[is.na(df_train$TEAM_PITCHING_SO)] <- mean(df_train$TEAM_PITCHING_SO, na.rm=TRUE)
```

```{r}
df_status(df_train)
```

Outliers
```{r}

boxplot(df_train$TEAM_PITCHING_SO)
boxplot(df_train$TEAM_PITCHING_H)
boxplot(df_train$TEAM_PITCHING_BB)
hist(df_train$TEAM_FIELDING_E)

```

John Tukey invented the box-and-whisker plot in 1977 to display IQR values, he picked 1.5×IQR as the demarkation line for outliers. 

```{r}
# assign the TEAM_PITCHING_SO outliers into a vector
outliers_SO <- boxplot(df_train$TEAM_PITCHING_SO, plot=FALSE)$out
# removing TEAM_PITCHING_SO outliers
df_train <- df_train[-which(df_train$TEAM_PITCHING_SO %in% outliers_SO),]
# demonstrating outliers removed, compared to above
boxplot(df_train$TEAM_PITCHING_SO)

# assign the TEAM_PITCHING_H outlier values into a vector
outliers_H <- boxplot(df_train$TEAM_PITCHING_H, plot=FALSE)$out
# removing TEAM_PITCHING_H outliers
df_train <- df_train[-which(df_train$TEAM_PITCHING_H %in% outliers_H),]
# demonstrating outliers removed, compared to above
boxplot(df_train$TEAM_PITCHING_H)

# assign the TEAM_PITCHING_BB outlier values into a vector
outliers_BB <- boxplot(df_train$TEAM_PITCHING_BB, plot=FALSE)$out
# removing TEAM_PITCHING_BB outliers
df_train <- df_train[-which(df_train$TEAM_PITCHING_BB %in% outliers_BB),]
# demonstrating outliers removed, compared to above
boxplot(df_train$TEAM_PITCHING_BB)
```

```{r}

df_train$TEAM_FIELDING_E = log10(df_train$TEAM_FIELDING_E)
hist(df_train$TEAM_FIELDING_E)
```


Do mean imputation
Add missing data indicators for some variables (3 variables that have >10% missing values)

# Model Building

Sample models - note that I haven't dealt with the missing values yet, so stats below are not right.

Suggest organizing the models we show:
- Kitchen sink model: use all variables at once
- Do step-by-step variable elimination: at each step, drop the variable that is not signif; stop when all variables are signif at 5%
- Try models with transformed variables: (a) try log or sqrt transform for skewed variables; (b) try aggregating related variables

```{r}

lm0 <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_BB + TEAM_BATTING_HBP + 
              TEAM_BASERUN_SB + TEAM_BASERUN_CS)
(lm0sum <- summary(lm0))

lm1 <- lm(TARGET_WINS ~ I(TEAM_BATTING_H + TEAM_BATTING_BB + TEAM_BATTING_HBP) + 
              I(TEAM_BASERUN_SB - TEAM_BASERUN_CS))
(lm1sum <- summary(lm1))

lm2 <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_BB + TEAM_BASERUN_SB + 
              TEAM_FIELDING_E + TEAM_PITCHING_H + TEAM_PITCHING_BB + TEAM_PITCHING_SO)
(lm2sum <- summary(lm2))

lm3 <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BATTING_3B +
              TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BATTING_HBP)
(lm3sum <- summary(lm3))

lm3a <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_HR + TEAM_BATTING_BB + 
               TEAM_BATTING_SO)
(lm3asum <- summary(lm3a))

lm4 <- lm(TARGET_WINS ~ TEAM_BASERUN_SB + TEAM_BASERUN_CS + TEAM_FIELDING_E +
              TEAM_FIELDING_DP)
(lm4sum <- summary(lm4))

lm5 <- lm(TARGET_WINS ~ TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_BB +
              TEAM_PITCHING_SO)
(lm5sum <- summary(lm5))

lm6 <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_HR + TEAM_BATTING_BB + 
              TEAM_BATTING_SO + TEAM_BASERUN_SB + TEAM_FIELDING_E + 
              TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_BB)
(lm6sum <- summary(lm6))

lm7 <- lm(TARGET_WINS ~ I(TEAM_BATTING_H + TEAM_BATTING_BB) + TEAM_BATTING_HR + 
              I(TEAM_BASERUN_SB - TEAM_BASERUN_CS) + TEAM_FIELDING_E +
              I(TEAM_PITCHING_H + TEAM_PITCHING_BB) + TEAM_PITCHING_HR)
(lm7sum <- summary(lm7))

```


# Model Selection

We start by reviewing summary statistics for the various models.

```{r}
library(knitr)

models <- list(lm0, lm1, lm2, lm3, lm4, lm5, lm6, lm7)
modsums <- list(lm0sum, lm1sum, lm2sum, lm3sum, lm4sum, lm5sum, lm6sum, lm7sum)
nmod <- length(modsums)

nvar <- integer(nmod)
sigma <- numeric(nmod)
rsq <- numeric(nmod)
adj_rsq <- numeric(nmod)
fstat_p <- numeric(nmod)

for (j in 1:nmod) {
    nvar[j] <- modsums[[j]]$df[1]
    sigma[j] <- modsums[[j]]$sigma
    rsq[j] <- modsums[[j]]$r.squared
    adj_rsq[j] <- modsums[[j]]$adj.r.squared
    fstat_p[j] <- 1 - pf(modsums[[j]]$fstatistic[1], modsums[[j]]$fstatistic[2], 
                         modsums[[j]]$fstatistic[3])
}

modnames <- paste0("lm", 0:7)

eval <- data.frame(Model = modnames, 
                   N_Variables = nvar,
                   Resid_Sigma = sigma,
                   R_Squared = rsq,
                   Adj_R_Squared = adj_rsq,
                   F_P_Value = fstat_p)

kable(eval, digits = 3, align = 'c', caption = 'Summary Evaluation Statistics')
```

>>> NEED to include MSE, RMSE 

```{r}
str(df_test)
head(df_test)

```


    
detach(df_train)

```

